---
title: "Overview Paper Figures"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    theme: spacelab
params:
  mode: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = TRUE, warning = FALSE, message = FALSE, tidy = TRUE, cache = FALSE)
knitr::opts_knit$set(root.dir = here::here())
```

# Setup

```{r library}
if(is.null(params$mode)) {
  library(knitr)
  library(targets)
  library(tarchetypes)
  library(ggplot2)
  library(tidyverse)
  library(ggh4x)
  source("packages.R")
  devtools::load_all(here())
}
```

```{r tar_load}
# config
tar_load(config)

# emf results data
tar_load(clean_data)
tar_load(clean_data_index)

# IPCC AR6 Data
tar_load(ipcc_ar6)

# figure maps
tar_load(figmap_op_band)
tar_load(figmap_op_cone)
tar_load(figmap_op_diffbar)
tar_load(figmap_op_stackbar)
tar_load(figmap_op_timeseries)
```

```{r supplementary data}
data_final = make_supplementary_data(clean_data) %>%
  filter(model %in% config$models_op & 
           scenario %in% config$sm_scenarios & 
           region == "United States")

data_2050 = data_final %>% filter(year == 2050)

data_w_aeo = clean_data %>%
  filter(model %in% config$models_op_aeo & 
           scenario %in% config$sm_scenarios & 
           region == "United States")

nontref = data_final %>% filter(scenario != "NT.Ref")
```

# Overview Paper Figures

```{r}
fig_no = 0
sm_fig_no = 0
```

## Emissions

### Net Zero Scenario CO2 Emissions, 2050

```{r}
fig_no = fig_no+1
fig_no

ref2020 = data_final %>%
  filter(year == 2020 & scenario == "NT.Ref" & variable == "Emissions|CO2|Energy and Industrial Processes") %>%
  mutate(label = "2020 Reference",
         scenario = "Reference") %>%
  select(model,scenario,year,value,label) %>%
  mutate(scenario = factor(scenario, levels = c("Reference", "Net Zero", "Net Zero+")))

fig <-
  print_graph(
    "stacked_bar",config,data_final,figmap_op_stackbar,7,"United States",
    scenario_rename = TRUE,
    level_var = c("Industrial Processes","Other","BECCS","Biofeedstock","DAC","Energy","LULUCF"),
    level_scen = c("Reference", "Net Zero", "Net Zero+")) +
  labs(title = "",
       y = expression(bold("Mt CO"[2]*"/yr"))) +
  geom_point(data = ref2020, aes(x = model, y = value, shape = label)) +
  theme(axis.text.x = element_text(vjust = .51),
        axis.title.y = element_text(face = "bold.italic")) +
  geom_hline(yintercept = 0) 
fig
ggsave(paste("output/round2/op/finals/fig",fig_no,".png",sep=""), width = 10, height = 6)
```

Statistics:

```{r}
reduction = data_2050 %>%
  filter(variable == "Emissions|CO2|Energy and Industrial Processes|Percent difference from 2020")

ref_red = reduction %>% filter(scenario == "NT.Ref") %>% mutate(value = round(value,1))
nz_ref = reduction %>% filter(scenario == "0by50.Ref") %>% mutate(value = round(value,1))
nz_adv = reduction %>% filter(scenario == "0by50.Adv") %>% mutate(value = round(value,1))

endog_lulucf = data_final %>%
  filter(variable == "Carbon Sequestration|LULUCF" & value != -800)

ind_process = unique((data_final %>% filter(variable == "Emissions|CO2|Industrial Processes"))$model)
no_ind_process = config$models_op[!config$models_op %in% ind_process]
```

Range of gross emissions reductions is `r max(ref_red$value)`-`r min(ref_red$value)`% in NT.Ref, `r max(nz_ref$value)`-`r min(nz_ref$value)`% in 0by50.Ref, and `r max(nz_adv$value)`-`r min(nz_adv$value)`%, in 0by50.Adv. Models that did not include the exogenous -800 Mt CO2/yr assumption for LULUCF include `r unique(endog_lulucf$model)`. Models that did not report industrial process emissions: `r no_ind_process`

### Sectoral Emissions, Difference from No Target Reference, 2050

```{r}
fig_no = fig_no+1 
fig_no

subpalettes = create_subpalettes(figmap_op_stackbar, config)
a_df = data_from_graph("stacked_bar",config,nontref,figmap_op_stackbar,9,"United States",
                           scenario_rename = TRUE,
                           level_var = c("Other","Petroleum Refining","Hydrogen","Buildings","Industry","Transportation","Electricity")) %>%
  ungroup() %>%
  select(model,scenario,region,unit,value,variable_rename) %>%
  mutate(facet = "Residual Emissions")
b_df = data_from_graph("diff_bar",config,data_final,figmap_op_diffbar,100,"United States",
                           scenario_rename = TRUE,
                           level_var = c("Other","Transportation","Industry","Buildings","Petroleum Refining","Hydrogen","Electricity")) %>%
  ungroup() %>%
  select(model,scenario,region,unit,diff,variable_rename) %>%
  rename(value = diff) %>%
  mutate(facet = "Difference from Reference")
df = rbind(a_df, b_df) %>%
  mutate(facet = factor(facet, levels = c("Residual Emissions","Difference from Reference")))

fig <- ggplot() +
  geom_bar(data = df,
           aes(x = model, y = value, fill = variable_rename),
           stat = "identity") +
  facet_grid(facet ~ scenario, scales = "free_y") +
  scale_subpalette(subpalettes, "Sector CO2 Emissions Only") +
  labs(y = expression(bold("Mt CO"[2]*"/yr")), x = "") +
  theme_emf() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .51)) +
  geom_hline(yintercept = 0, alpha = 0.5, color = "black")
fig
ggsave(paste("output/round2/op/finals/fig",fig_no,".png",sep=""), width = 9, height = 5)
```

Statistics:

```{r}
diff = data_2050 %>%
  filter(scenario != "NT.Ref") %>% 
  mutate(value = round(value,0))
 
elcdiff2020 = diff %>% filter(variable == "Emissions|CO2|Energy|Supply|Electricity|Percent difference from 2020")  %>%
  select(model,scenario,unit,year,value)
elcdiffntref = diff %>% filter(variable == "Emissions|CO2|Energy|Supply|Electricity|Percent Difference from NT.Ref")  %>%
  select(model,scenario,unit,year,value)

blddiffntref = diff %>% filter(variable == "Emissions|CO2|Energy|Demand|Buildings|Percent Difference from NT.Ref") %>%
  select(model,scenario,unit,year,value)
inddiffntref = diff %>% filter(variable == "Emissions|CO2|Energy|Demand|Industry|Percent Difference from NT.Ref") %>%
  select(model,scenario,unit,year,value)
trndiffntref = diff %>% filter(variable == "Emissions|CO2|Energy|Demand|Transportation|Percent Difference from NT.Ref") %>%
  select(model,scenario,unit,year,value)
```

*** NEED TO REVISIT SOME OF THESE STATS - DONT MATCH UP TO INDEXED VALUES AND NEED VERBAGE FOR DESCRIPTION ****

Net zero scenario power sector emissions are reduced by no less than `r min(elcdiff2020$value)`% from 2020 to 2050, ranging from `r min(elcdiffntref$value)`-`r max(elcdiffntref$value)`% reductions over 2050 emissions in the no target reference scenario. Proportionate to reference scenario 2050 emissions, industrial emissions reduce between `r min(inddiffntref$value)`-`r max(inddiffntref$value)`%, buildings emissions reduce between `r min(blddiffntref$value)`-`r max(blddiffntref$value)`%, and transportation emissions reduce between `r min(trndiffntref$value)`-`r max(trndiffntref$value)`%.

### Carbon Sequestration, 2050

```{r}
fig_no = fig_no+1 
fig_no

fig <- print_graph("stacked_bar",config,data_final,figmap_op_stackbar,3,"United States",
                    scenario_rename = TRUE,
                    level_var = c("Fossil CCS","BECCS","DAC","LULUCF")) +
  labs(title = "",
       y = expression(bold("Mt CO"[2]*"/yr"))) +
  theme(axis.text.x = element_text(vjust = .51))
fig
ggsave(paste("output/round2/op/finals/fig",fig_no,".png",sep=""), width = 9, height = 5)
```

## Carbon Price

```{r}
fig_no = fig_no+1 
fig_no

fig <- print_graph("time_series",config,data_final,figmap_op_timeseries,8,"United States",
                    scenario_rename = TRUE) +
  labs(title = "",
       y = expression(bold("2018$USD/t CO"[2])))
fig
ggsave(paste("output/round2/op/finals/fig",fig_no,".png",sep=""), width = 8, height = 4)
```

Statistics:

```{r}
price_diff <- data_2050 %>% 
  filter(variable == "Price|Carbon|Difference from 0by50.Ref" & scenario == "0by50.Adv") %>%
  mutate(value = round(value*-1,1))

# TODO: which models didnt submit carbon prices?
```

Carbon prices are reduced by `r min(price_diff$value)`-`r max(price_diff$value)`% with the advanced technology assumptions

## Final Energy

### Final Energy, Difference from NT.Ref

```{r}
fig_no = fig_no+1 
fig_no

subpalettes = create_subpalettes(figmap_op_diffbar, config)
level_var = c("Synthetic Gas","Synthetic Liquids","Biogas","Biomass Liquids","Biomass Solids","Coal","Gas","Oil","Hydrogen","Electricity")
level_scen = c("Reference", "Net Zero", "Net Zero+")

diff_param = list(
  plot_type = "diff_bar", config = config, emf_data_long = data_final, figmap = figmap_op_diffbar,
  figure_num = 14, reg = "United States", scenario_rename = TRUE, level_var = level_var, level_scen = level_scen)

ref_param = list(
  plot_type = "stacked_bar", config = config, emf_data_long = data_final, figmap = figmap_op_stackbar,
  figure_num = 6, reg = "United States", scenario_rename = TRUE, level_var = level_var, level_scen = level_scen)

diff_df = do.call("data_from_graph", diff_param) %>%
  ungroup() %>%
  select(model,scenario,region,year,unit,diff,variable_rename) %>%
  rename(value = diff) %>%
  mutate(facet = case_when(
    scenario == "Net Zero" ~ "Difference: Net Zero",
    scenario == "Net Zero+" ~ "Difference: Net Zero+",
    TRUE~"fix"))

ref_df = do.call("data_from_graph", ref_param)  %>%
  filter(scenario == "Reference") %>%
  ungroup() %>%
  select(model,scenario,region,unit,year,value,variable_rename) %>%
  mutate(facet = "Reference")

df = rbind(diff_df, ref_df) %>%
  filter(year != 2020) %>%
  mutate(facet = factor(facet, levels = c("Reference","Difference: Net Zero","Difference: Net Zero+")))

hline_adv = df %>%
  filter(facet == "Difference: Net Zero+" & !model %in% c("ADAGE","AnyMOD","EC-MSMR","EP-RIO","FECM-NEMS","gTech","TEMOA","US-REGEN","USREP-ReEDS"))
hline_ref = df %>%
  filter(facet == "Difference: Net Zero")
hline = rbind(hline_adv,hline_ref)

fig <- ggplot() +
  geom_bar(data = df,
           aes(x = year, y = value, fill = variable_rename),
           stat = "identity") +
  facet_grid(facet ~ model, scales = "free_y") +
  facetted_pos_scales(
    y = list(
      facet == "Reference" ~ scale_y_continuous(limits = c(0, 100)),
      facet == "Difference: Net Zero" ~ scale_y_continuous(limits = c(-50, 50)),
      facet == "Difference: Net Zero+" ~ scale_y_continuous(limits = c(-50, 50))
      )) +
  geom_hline(data = hline, aes(yintercept = 0), color="black", linewidth=.5, alpha=0.7) +
  scale_subpalette(subpalettes, "Final Energy") +
  labs(y = "EJ/yr", x = "") +
  theme_emf() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .51),
        strip.text.y = element_text(size = 10, color = "black",face = "bold")) +
  bottom1
fig
ggsave(paste("output/round2/op/finals/fig",fig_no,".png",sep=""), width = 20, height = 11)
```

```{r}
finalnrg_diff2020 = data_2050 %>%
  filter(variable == "Final Energy|Percent difference from 2020")

finalnrg_diffntref = data_2050 %>%
  filter(variable == "Final Energy|Percent difference from NT.Ref" &
           scenario != "NT.Ref")
```

### Sector Final Energy, Difference from NT.Ref, 2050

```{r}
fig_no = fig_no+1 
fig_no

df = data_from_graph("diff_bar",config,data_final,figmap_op_diffbar,12,"United States",
                          scenario_rename = TRUE,
                          level_scen = c("Reference","Net Zero","Net Zero+"),
                          level_var = c("Synthetic Gas","Synthetic Liquids","Biogas","Biomass Liquids","Biomass Solids",
                                             "Coal","Gas","Oil","Hydrogen","Electricity")) %>%
  mutate(sector = case_when(
    str_detect(variable, "Buildings") ~ "Buildings",
    str_detect(variable, "Industry") ~ "Industry",
    str_detect(variable, "Transportation") ~ "Transportation")) 

sub_palettes = create_subpalettes(figmap_op_diffbar, config)

fig = ggplot(df, aes(x = model, y = diff)) +
  geom_col(aes(fill = variable_rename)) +
  facet_grid(scenario~sector,
               space = "free_x", scales = "fixed") +
  labs(title = "",
         x = "",
         y = "EJ/yr",
         fill = "") +
  theme_emf() +
  theme(axis.text.x = element_text(angle = 90, hjust=1)) +
  scale_subpalette(sub_palettes, "Industry Final Energy Mix - Difference from NT.Ref") + 
  theme(axis.text.x = element_text(vjust = .51)) +
  geom_hline(yintercept = 0, alpha = 0.7, color = "black")
fig
ggsave(paste("output/round2/op/finals/fig",fig_no,".png",sep=""), width = 8, height = 6)
```

Statistics:

```{r}
subset = data_2050 %>%
  mutate(value = round(value * -1,1))

bld = subset %>% filter(variable == "Final Energy|Buildings|Percent difference from NT.Ref")
bld_gas = subset %>% filter(variable == "Final Energy|Buildings|Gas|Percent difference from NT.Ref")
ind_gas = subset %>% filter(variable == "Final Energy|Industry|Gas|Percent difference from NT.Ref")
```

There is strong consensus between models for buildings to transition away from gas, a `r min(bld$value)`-`r max(bld$value)`% reduction over reference, and significant total final energy demand decreases, a `r min(bld_gas$value)`-`r max(bld_gas$value)`% reduction over reference. All models show reductions in industrial gas demand, `r min(ind_gas$value)`-`r max(ind_gas$value)`% over the reference scenario. 

## Electrification

### Percent Elc of Final Energy, Time Series, Objective Values

```{r}
fig_no = fig_no+1 
fig_no

vars = c("Final Energy|Percent Electricity", "Final Energy|Buildings|Percent Electricity",
         "Final Energy|Industry|Percent Electricity","Final Energy|Transportation|Percent Electricity")

elc = data_2050 %>% filter(variable %in% vars)

ref_elc = clean_data %>%
  filter(variable %in% vars & year == 2020 & model == "AEO2021") %>%
  mutate(value = case_when(
    variable == "Final Energy|Buildings|Percent Electricity" ~ 0.47, # number from Chioke, our number is wrong
    variable == "Final Energy|Industry|Percent Electricity" ~ 0.13, # number calculated from AEO 2021, ~3 quads elc ~25 quads total
    TRUE~value))

elc_all = rbind(elc, ref_elc) %>%
  mutate(
    variable = case_when(
      variable == "Final Energy|Percent Electricity" ~ "Total",
      str_detect("Buildings", variable) ~ "Buildings",
      str_detect("Industry", variable) ~ "Industry",
      str_detect("Transportation", variable) ~ "Transportation",
      TRUE ~ variable
    ),
    scenario = case_when(
      scenario == "NT.Ref" ~ "Reference",
      scenario == "0by50.Ref" ~ "Net Zero",
      scenario == "0by50.Adv" ~ "Net Zero+",
      TRUE ~ scenario),
    year = case_when(
      year == 2020 ~ "2020 Reference",
      year == 2050 ~ "2050"
    )) %>%
  mutate(value = value * 100) %>%
  mutate(
    scenario = factor(scenario, levels = c("Reference", "Net Zero", "Net Zero+")),
    variable = factor(variable, levels = c("Total", "Transportation", "Industry", "Buildings")))

means = elc_all %>%
  filter(year == 2050) %>%
  group_by(variable,scenario) %>%
  summarise(value = mean(value)) %>%
  mutate(year = "2050 Mean")

ref_points = elc_all %>% filter(year == "2020 Reference")
mod_points = elc_all %>% filter(year == "2050")
  
sub_palettes = create_subpalettes(figmap_op_timeseries, config)
shapes <- c("2020 Reference"=18, "2050"=16, "2050 Mean"=17)

fig <- ggplot() +
  geom_point(data=mod_points, aes(x=scenario,y=value, color=model, shape = year)) +
  geom_line(data=mod_points, aes(x=scenario,y=value, group = model, color=model), linewidth = 1, alpha = .85) +
  geom_point(data=ref_points, aes(x=scenario,y=value, shape = year), size = 4, color = "black") +
  geom_point(data = means, aes(x=scenario,y=value, shape = year), size = 3, color = "black") +
  facet_grid(~variable) +
  labs(title = "",
       x="", y = "Electricity Share in Final Energy (%)") +
  theme_emf() +
  scale_subpalette(sub_palettes, "Percent Electricity") +
  scale_shape_manual(values = shapes) +
  theme(axis.text.x = element_text(angle = 90, hjust = .51)) +
  guides(color = guide_legend(ncol=1))
fig
ggsave(paste("output/round2/op/finals/fig",fig_no,".png",sep=""), width = 10, height = 6)
```

Statistics:

```{r}
subset = data_2050 %>% 
  filter(scenario != "NT.Ref") %>%
  mutate(value = round(value * 100,1))
elcbld = subset %>% filter(variable == "Final Energy|Buildings|Percent Electricity")
elcind = subset %>% filter(variable == "Final Energy|Industry|Percent Electricity")
elctrn = subset %>% filter(variable == "Final Energy|Transportation|Percent Electricity")
elctot = subset %>% filter(variable == "Final Energy|Percent Electricity")
```

Across net zero scenarios, there is a much steeper increase in that proportion from 2020 to 2050, `r min(elcbld$value)`-`r max(elcbld$value)`% for buildings, `r min(elcind$value)`-`r max(elcind$value)`% in industry, and `r min(elctrn$value)`-`r max(elctrn$value)`% in transportation, leading to a total increase of `r min(elctot$value)`-`r max(elctot$value)`%. 

### Generation Mix, Difference from NT.Ref

```{r}
fig_no = fig_no+1 
fig_no

nogtech <- clean_data %>% 
  filter(model != "gTech" & year != 2020) %>%
  mutate(label = "Net")

param = list(
  plot_type = "diff_bar", 
  config = config, 
  emf_data_long = nogtech, 
  figmap = figmap_op_diffbar,
  figure_num = 7, 
  reg = "United States", 
  scenario_rename = TRUE,
  level_var = c("Wind","Solar","Hydrogen","Hydro","Biomass","Coal","Gas","Oil","Nuclear","Other"))

hline_adv = do.call("data_from_graph",param) %>%
  filter(scenario == "Net Zero+" & !model %in% c("ADAGE","EC-MSMR","FECM-NEMS","TEMOA","USREP-ReEDS"))
hline_ref = do.call("data_from_graph",param) %>%
  filter(scenario == "Net Zero")
hline = rbind(hline_adv,hline_ref)

fig <- do.call("print_graph",param) + 
  labs(title = "",
       x="") + 
  geom_hline(data = hline, aes(yintercept = 0), color="black", linewidth=.75, alpha=0.7) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .51),
        strip.text.x = element_text(size = 10, color = "black",face = "bold"),
        strip.text.y = element_text(size = 11, color = "black",face = "bold")) +
  bottom1
fig
ggsave(paste("output/round2/op/finals/fig",fig_no,".png",sep=""), width = 16, height = 8)
```

## Net-Zero Pathways

### Literature Comparison

```{r}
# EMF <--> Comparison variable and axis mapping
var_map = readxl::read_xlsx("data-raw/extra-data/bistline_var_map.xlsx")

# Read in comparison data
comp = readxl::read_xlsx("data-raw/extra-data/bistline_scatter.xlsx") %>%
  pivot_longer(cols = `elsr (% final)`:`ffr (% 2015)`, names_to = "bistline_var", values_to = "value") %>%
  left_join(var_map, by = "bistline_var", relationship = "many-to-many") %>%
  filter(!Region %in% c("China","Japan","Mexico")) %>%
  mutate(
    `CO2 Target` = case_when(
      `CO2 Target` == "DD" ~ "Deep Decarbonization",
      `CO2 Target` == "NZ" ~ "Net Zero",
      `CO2 Target` == "REF" ~ "Reference",
      `CO2 Target` == "BASE" ~ "2015"),
    Region = case_when(
      Region == "OECD" ~ "OECD: IPCC SR15",
      Region == "United States" ~ "US: EMF 37",
      TRUE~Region), 
    Region = case_when(
      !Region %in% c("OECD: IPCC SR15", "US: EMF 37") ~ paste(Region,"*",sep=""),
      TRUE~Region)
    )

# color, shape, alpha palettes
var_palette = c(unique(comp$Region),"US: IPCC AR6")
names(var_palette) = var_palette
sub_palettes[["comparisons"]] = var_palette
sub_palettes = sub_palettes %>%
  map(~find_color(.x, color_map))

shapes <- c("2015" = 4, "Deep Decarbonization" = 17, "Net Zero" = 16, "Reference" = 1)
```

```{r}
df_comp = comp %>% 
  filter(figure == 1 & !Region %in% c("Asia*","EU*","Latin America*")) %>% 
  select(-bistline_var,-figure,-axis_title,-bistline_var_no,-emf_var) %>%
  mutate(value = value*100) %>%
  pivot_wider(names_from = axis, values_from = value) %>%
  mutate(source = "Literature") %>%
  select(`CO2 Target`,Region,x,y,source)

df_emf = data_2050 %>% 
  filter(variable %in% c("Final Energy|Percent Electricity","Final Energy|Electricity|Percent of 2015")) %>%
  select(model,scenario,region,value,variable) %>%
  filter(value != Inf) %>%
  pivot_wider(names_from = "variable", values_from = "value") %>%
  rename(x = `Final Energy|Percent Electricity`,
         y = `Final Energy|Electricity|Percent of 2015`,
         Region = region) %>%
  mutate(
    `CO2 Target` = case_when(scenario == "NT.Ref" ~ "Reference",
                             scenario != "NT.Ref" ~ "Net Zero"),
    source = "EMF37",
    x = x * 100,
    Region = "US: EMF 37"
  ) %>%
  select(`CO2 Target`,Region,x,y,source)

fig_a = ggplot() + 
  geom_point(data = ipcc_ar6$ipcc_a , aes(x = x, y = y, color = Region, shape = `CO2 Target`), size = 2.4, alpha = 0.5) + 
  geom_point(data = df_comp, aes(x = x, y = y, color = Region, shape = `CO2 Target`), size = 2.4, alpha = 0.4) + 
  geom_point(data = df_emf, aes(x = x, y = y, color = Region, shape = `CO2 Target`), size = 3) + 
  scale_subpalette(sub_palettes, "comparisons") +
  theme_emf() +
  theme(axis.title.x = element_text(face="bold")) +
  labs(y = "Electricity Demand (% 2015)",
       x = "Electricity Share in Final Energy (%)",
       title = "a.") +
  scale_shape_manual(values = shapes) +
  nolegend
fig_a
```

```{r}
df_comp = comp %>% 
  filter(figure == 4 & !Region %in% c("Asia*","EU*","Latin America*")) %>% 
  # filter(`CO2 Target` != "2015") %>%
  filter(!(Region == "Asia*" & `CO2 Target` == "Reference")) %>%
  select(-bistline_var,-figure,-axis_title,-bistline_var_no,-emf_var) %>%
  mutate(value = value * 100) %>%
  pivot_wider(names_from = axis, values_from = value) %>%
  mutate(source = "Literature") %>%
  select(`CO2 Target`,Region,x,y,source)

df_emf = data_2050 %>% 
  filter(variable %in% c("Primary Energy|Fossil|Percent difference from 2015","Emissions|CO2|Percent difference from 2015")) %>%
  filter(model != "ADAGE") %>%
  select(model,scenario,region,value,variable) %>%
  mutate(value = value*-1) %>%
  filter(value != -Inf) %>%
  pivot_wider(names_from = "variable", values_from = "value") %>%
  rename(x = `Primary Energy|Fossil|Percent difference from 2015`,
         y = `Emissions|CO2|Percent difference from 2015`,
         Region = region) %>%
  mutate(`CO2 Target` = case_when(scenario == "NT.Ref" ~ "Reference",
                                  scenario != "NT.Ref" ~ "Net Zero"),
         source = "EMF37",
    Region = "US: EMF 37") %>%
  select(`CO2 Target`,Region,x,y,source)

# TODO: IPCC emissions and fossil fuel reductions still need to be figured out
test = ipcc_ar6$ipcc_b %>%
  filter(x > 30 & y > 30 & x < 100 & y < 100)

fig_b = ggplot() + 
  geom_point(data = test, aes(x = x, y = y, color = Region, shape = `CO2 Target`), size = 2.4, alpha = 0.4) + 
  geom_point(data = df_comp, aes(x = x, y = y, color = Region, shape = `CO2 Target`), size = 2.4, alpha = 0.4) + 
  geom_point(data = df_emf, aes(x = x, y = y, color = Region, shape = `CO2 Target`), size = 3) + 
  scale_subpalette(sub_palettes, "comparisons") +
  theme_emf() +
  theme(axis.title.x = element_text(face="bold")) +
  scale_shape_manual(values = shapes) +
  labs(x = "Fossil Fuel Reduction from 2015 (%)",
       # y = "Gross Economy CO2 Reductions from 2015 (%)",
       y = expression(bold("Gross Economy CO"[2]*" Reductions from 2015 (%)")),
       title = "b.")
fig_b
```

```{r}
fig_no = fig_no+1 
fig_no
both = fig_a + fig_b + plot_layout(guides = 'collect')
both
ggsave(paste("output/round2/op/finals/fig",fig_no,".png",sep=""), width = 12, height = 5)
```

### Energy and Emissions Summary w/ Carbon Price

```{r}
fig_no = fig_no+1 
fig_no

co2price = data_2050 %>% 
  filter(scenario == "0by50.Ref" & variable == "Price|Carbon") %>% 
  mutate(variable = "CO2 Emissions",
         # label = paste("$",round(value,0),sep=""),
         label = round(value,0),
         legend = "CO2 Price (2018$USD/ton)")

fig <- print_graph("stacked_bar",config,nontref,figmap_op_stackbar,15.1,"United States",
                   level_var = c("CO2 Emissions","Fossil Energy Demand","Total Energy Demand","Electricity Demand")) +
  #facet_grid(.~model) +
  facet_wrap(~model, nrow = 2) +
  geom_point(data = co2price, aes(x=variable,y=value/600, shape = legend), color = "#2A4849") +
  geom_text_repel(data = co2price, 
            aes(x=variable,y=value/600, label = label), 
            nudge_x = .15, nudge_y = .15, color = "#2A4849", size = 4, fontface = "bold") +
  bottom1 +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        strip.text.x = element_text(size = 9)) +
  labs(y = "Annual Rate of Change 2020-2050 (%)", title = "")
fig
ggsave(paste("output/round2/op/finals/fig",fig_no,".png",sep=""), width = 9, height = 5)
```

# Supplementary Materials

## Calibration

### GDP 

```{r}
sm_fig_no = sm_fig_no+1
sm_fig_no

gdp_df = data_w_aeo %>% filter(!model %in% c("NATEM","gTech"))

fig_a = print_graph("time_series",config,gdp_df,figmap_op_timeseries,12.1,"United States",
                     scenario_rename = TRUE) +
  labs(x="",title = "a.",y = "Billion 2018$USD/yr")
fig_a
fig_b = print_graph("time_series",config,gdp_df,figmap_op_timeseries,12.2,"United States",
                     scenario_rename = TRUE) +
  labs(title = "b.",
       y = "GDP % Difference from AEO 2022",
       x="")
fig_b
both = fig_a + fig_b + plot_layout(guides = 'collect')
both
ggsave(paste("output/round2/op/finals/supmat/fig",sm_fig_no,".png",sep=""), width = 8, height = 4)
```

### Population

```{r}
sm_fig_no = sm_fig_no+1
sm_fig_no
fig = print_graph("time_series",config,data_w_aeo,figmap_op_timeseries,13,"United States",
                     scenario_rename = TRUE) +
  labs(x="",title = "",y="Million People")
fig
ggsave(paste("output/round2/op/finals/supmat/fig",sm_fig_no,".png",sep=""), width = 6, height = 4)
```

## Emissions

### Sectoral Emissions

#### By Model, Absolute Values

```{r}
sm_fig_no = sm_fig_no+1
sm_fig_no
fig = print_graph("stacked_bar",config,data_final,figmap_op_stackbar,8.1,"United States",
                     scenario_rename = TRUE,
                  level_scen = c("Reference","Net Zero","Net Zero+")) +
  labs(x="",title = "",y = expression(bold("Mt CO"[2]*"/yr"))) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .51),
        strip.text.y = element_text(size = 10, color = "black",face = "bold"),
        strip.text.x = element_text(size = 8, color = "black",face = "bold")) +
  bottom1
fig
ggsave(paste("output/round2/op/finals/supmat/fig",sm_fig_no,".png",sep=""), width = 15, height = 5)
```

#### By Model, Difference from Net Zero

```{r}
sm_fig_no = sm_fig_no+1
sm_fig_no

subpalettes = create_subpalettes(figmap_op_diffbar, config)

df = data_from_graph("diff_bar",config,nontref,figmap_op_diffbar,100.1,"United States") %>%
  ungroup() %>%
  select(model,scenario,region,unit,diff,variable_rename) %>%
  rename(value = diff) %>%
  mutate(facet = "Net Zero +: Difference from Net Zero")

fig <- ggplot() +
  geom_bar(data = df,
           aes(x = model, y = value, fill = variable_rename),
           stat = "identity") +
  facet_grid(. ~ facet, scales = "free_y") +
  scale_subpalette(subpalettes, "Sector CO2 Emissions Only- Difference from 0by50.Ref") +
  labs(y = expression(bold("Mt CO"[2]*"/yr")), x = "") +
  theme_emf() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .51)) +
  geom_hline(yintercept = 0, alpha = 0.5, color = "black")
fig
ggsave(paste("output/round2/op/finals/supmat/fig",sm_fig_no,".png",sep=""), width = 5, height = 4)
```

### Carbon Capture and Removals

#### Total

```{r}
sm_fig_no = sm_fig_no+1
sm_fig_no
fig = print_graph("stacked_bar",config,data_final,figmap_op_stackbar,3.1,"United States",
                     scenario_rename = TRUE) +
  labs(x="",title = "",y = expression(bold("Mt CO"[2]*"/yr"))) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .51),
        strip.text.y = element_text(size = 10, color = "black",face = "bold"),
        strip.text.x = element_text(size = 8, color = "black",face = "bold")) +
  bottom1
fig
ggsave(paste("output/round2/op/finals/supmat/fig",sm_fig_no,".png",sep=""), width = 15, height = 5)
```

#### Carbon Capture Accounting

```{r}
sm_fig_no = sm_fig_no+1
sm_fig_no

df = data_from_graph("stacked_bar",config,nontref,figmap_op_stackbar,102,"United States",
                     scenario_rename = TRUE,
                     level_scen = c("Reference","Net Zero","Net Zero+")) %>%
  filter(year == 2050) %>%
  mutate(capture = case_when(
    str_detect(variable, "Biomass") ~ "BECCS",
    TRUE~"Fossil CCS"
  ))

sub_palettes = create_subpalettes(figmap_op_stackbar, config)

fig = ggplot(df, aes(x = model, y = value)) +
  geom_col(aes(fill = variable_rename)) +
  facet_grid(scenario~capture,
               space = "free_x", scales = "fixed") +
  theme_emf() +
  theme(axis.text.x = element_text(angle = 90, hjust=1)) +
  scale_subpalette(sub_palettes, "CCS Accounting") + 
  theme(axis.text.x = element_text(vjust = .51)) +
  geom_hline(yintercept = 0, alpha = 0.7, color = "black") +
  labs(y = expression(bold("Captured Carbon (Mt CO"[2]*"/yr)")),
       title="",
       x="")
fig
ggsave(paste("output/round2/op/finals/supmat/fig",sm_fig_no,".png",sep=""), width = 9, height = 5)
```

#### Power Sector Emissions and Carbon Capture

```{r}
sm_fig_no = sm_fig_no+1
sm_fig_no
fig <- print_graph("stacked_bar",config,nontref,figmap_op_stackbar,100,"United States",
                     scenario_rename = TRUE,
                   level_scen = c("Reference","Net Zero","Net Zero+")) +
  labs(y = expression(bold("Power Sector Emissions and CCS (Mt CO"[2]*"/yr)")),title="") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .51),
        strip.text.y = element_text(size = 10, color = "black",face = "bold"),
        strip.text.x = element_text(size = 8, color = "black",face = "bold")) +
  bottom1
fig
ggsave(paste("output/round2/op/finals/supmat/fig",sm_fig_no,".png",sep=""), width = 15, height = 5)
```

#### Industry Emissions and Carbon Capture

```{r}
sm_fig_no = sm_fig_no+1
sm_fig_no

norio = nontref %>% filter(model != "EP-RIO")

fig <- print_graph("stacked_bar",config,norio,figmap_op_stackbar,101,"United States",
                     scenario_rename = TRUE,
                   level_scen = c("Reference","Net Zero","Net Zero+")) +
  labs(y = expression(bold("Industrial Emissions and CCS (Mt CO"[2]*"/yr)")),title="") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .51),
        strip.text.y = element_text(size = 10, color = "black",face = "bold"),
        strip.text.x = element_text(size = 8, color = "black",face = "bold")) +
  bottom1
fig
ggsave(paste("output/round2/op/finals/supmat/fig",sm_fig_no,".png",sep=""), width = 15, height = 5)
```

## Final Energy

### Absolute Values, Tme Series

```{r}
sm_fig_no = sm_fig_no+1
sm_fig_no
fig = print_graph("stacked_bar",config,data_final,figmap_op_stackbar,6.5,"United States",
                   scenario_rename = TRUE,
                   level_scen = c("Reference", "Net Zero", "Net Zero+"),
                   level_var = c("Synthetic Gas","Synthetic Liquids","Biogas","Biomass Liquids","Biomass Solids","Coal","Gas","Oil","Hydrogen","Electricity")) +
  labs(title = "",y="Final Energy (EJ/yr)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .51),
        strip.text.y = element_text(size = 10, color = "black",face = "bold"),
        strip.text.x = element_text(size = 8, color = "black",face = "bold")) +
  bottom1
fig
ggsave(paste("output/round2/op/finals/supmat/fig",sm_fig_no,".png",sep=""), width = 15, height = 5)
```

Statistics:

```{r}
#TODO: Need to rework all of these stats, changed with updated results and inclusion of more models
#TODO: add stats for annual growth rate of electricity to compare how quickly it grows in NT.Ref vs 0by50 scenarios

finalnrg2020 = data_2050 %>%
  filter(variable == "Final Energy|Percent difference from 2020") %>%
  mutate(value = round(value,1))

diff2020ntref = finalnrg2020 %>% filter(scenario == "NT.Ref" & model != "US-REGEN")
diff2020ntref_usregen = finalnrg2020 %>% filter(scenario == "NT.Ref") %>% filter(model == "US-REGEN")
diff2020nzref = finalnrg2020 %>% filter(scenario == "0by50.Ref" & !model %in% c("ADAGE", "EC-MSMR")) %>% mutate(value = value * -1)
diff2020nzref_adage = finalnrg2020 %>% filter(scenario == "0by50.Ref" & model == "ADAGE")
diff2020nzref_ecmsmr = finalnrg2020 %>% filter(scenario == "0by50.Ref" & model == "EC-MSMR")
diff2020nzadv = finalnrg2020 %>% filter(scenario == "0by50.Adv" & model != "GCAM")
diff2020nzadv_gcam = finalnrg2020 %>% filter(scenario == "0by50.Adv" & model == "GCAM")

finalnrgntref = data_2050 %>%
  filter(variable == "Final Energy|Percent difference from NT.Ref") %>%
  mutate(value = round(value * -1,1))

diffntref_nzref = finalnrgntref %>% filter(scenario == "0by50.Ref" & model != "ADAGE")
diffntref_nzref_adage = finalnrgntref %>% filter(scenario == "0by50.Ref" & model == "ADAGE")
diffntref_nzadv = finalnrgntref %>% filter(scenario == "0by50.Adv")

fuels = data_2050 %>%
  mutate(value = round((value * -1)/30,1))

oilnrg = fuels %>% filter(variable == "Final Energy|Oil|Percent difference from 2020" & scenario == "0by50.Ref") 

gasnrg_all = fuels %>% filter(variable == "Final Energy|Gas|Percent difference from 2020" & scenario == "0by50.Ref")
gasnrg_sm = gasnrg_all %>% filter(!model %in% c("ADAGE","gTech"))
gasnrg_adage = gasnrg_all %>% filter(model == "ADAGE")
gasnrg_gtech = gasnrg_all %>% filter(model == "gTech")

elcnrg = data_2050 %>%
  filter(variable == "Final Energy|Electricity|Percent difference from 2020") %>%
  mutate(value = round((value)/30,1))
elcnrgnzref = elcnrg %>% filter(scenario == "0by50.Ref")
elcnrgntref = elcnrg %>% filter(scenario == "NT.Ref")
```

Final energy demand increases in the reference scenario from 2020 to 2050 for all by one model by `r min(diff2020ntref$value)`-`r max(diff2020ntref$value)`% (US-REGEN, `r diff2020ntref_usregen$value`%) and decreases for most models in the net zero reference and advanced scenarios, `r max(diff2020nzref$value)`-`r min(diff2020nzref$value)`% (except ADAGE, +`r diff2020nzref_adage$value`%, and EC-MSMR, +`r diff2020nzref_ecmsmr$value`%) and `r max(diff2020nzadv$value)`-`r min(diff2020nzadv$value)`% (except GCAM, `r diff2020nzadv_gcam$value`%) respectively. Total final energy demand in 2050 decreases from the reference to the net zero reference scenario in most models by `r min(diffntref_nzref$value)`-`r max(diffntref_nzref$value)`% (except ADAGE, `r diffntref_nzref_adage$value`%), and a similar range to the net zero advanced scenario, `r min(diffntref_nzadv$value)`-`r max(diffntref_nzadv$value)`%. Final demand for gas and oil consumption falls from 2020 to 2050 in the net zero reference scenario for almost all models, by `r min(gasnrg_sm$value)`-`r max(gasnrg_sm$value)`% per year (except ADAGE, `r gasnrg_adage$value`%, and gTech, `r gasnrg_gtech$value`%) and `r min(oilnrg$value)`-`r max(oilnrg$value)`% per year, respectively. Electricity demand in the net zero reference scenarios grows slightly faster, `r min(elcnrgnzref$value)`-`r max(elcnrgnzref$value)`% per year, than in the no target reference, `r min(elcnrgntref$value)`-`r max(elcnrgntref$value)`% per year.

### Buildings

```{r}
sm_fig_no = sm_fig_no+1
sm_fig_no
fig <- print_graph("stacked_bar",config,data_final,figmap_op_stackbar,23,"United States",
                     scenario_rename = TRUE,
                   level_scen = c("Reference","Net Zero","Net Zero+")) +
  labs(title="",x="",y="Buildings Final Energy (EJ/yr)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .51),
        strip.text.y = element_text(size = 10, color = "black",face = "bold"),
        strip.text.x = element_text(size = 8, color = "black",face = "bold")) +
  bottom1
fig
ggsave(paste("output/round2/op/finals/supmat/fig",sm_fig_no,".png",sep=""), width = 15, height = 5)
```

### Industry

```{r}
sm_fig_no = sm_fig_no+1
sm_fig_no
fig <- print_graph("stacked_bar",config,data_final,figmap_op_stackbar,24,"United States",
                     scenario_rename = TRUE,
                   level_scen = c("Reference","Net Zero","Net Zero+")) +
  labs(title="",x="",y="Industry Final Energy (EJ/yr)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .51),
        strip.text.y = element_text(size = 10, color = "black",face = "bold"),
        strip.text.x = element_text(size = 8, color = "black",face = "bold")) +
  bottom1
fig
ggsave(paste("output/round2/op/finals/supmat/fig",sm_fig_no,".png",sep=""), width = 15, height = 5)
```

### Transportation

```{r}
sm_fig_no = sm_fig_no+1
sm_fig_no
fig <- print_graph("stacked_bar",config,data_final,figmap_op_stackbar,25,"United States",
                     scenario_rename = TRUE,
                   level_scen = c("Reference","Net Zero","Net Zero+")) +
  labs(title="",x="",y="Transportation Final Energy (EJ/yr)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .51),
        strip.text.y = element_text(size = 10, color = "black",face = "bold"),
        strip.text.x = element_text(size = 8, color = "black",face = "bold")) +
  bottom1
fig
ggsave(paste("output/round2/op/finals/supmat/fig",sm_fig_no,".png",sep=""), width = 15, height = 5)
```

## Electricity Generation

### Totals

```{r}
sm_fig_no = sm_fig_no+1
sm_fig_no
nogtech = data_final %>% filter(model != "gTech")
fig <- print_graph("stacked_bar",config,nogtech,figmap_op_stackbar,20,"United States",
                     scenario_rename = TRUE,
                   level_scen = c("Reference","Net Zero","Net Zero+")) +
  labs(title = "") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .51),
        strip.text.y = element_text(size = 10, color = "black",face = "bold"),
        strip.text.x = element_text(size = 8, color = "black",face = "bold")) +
  bottom1
fig
ggsave(paste("output/round2/op/finals/supmat/fig",sm_fig_no,".png",sep=""), width = 15, height = 5)
```

### Combustion w/ and w/o CCS

```{r}
sm_fig_no = sm_fig_no+1
sm_fig_no
noecmsmr = data_final %>% filter(!model %in% c("EC-MSMR","EPS"))
fig <- print_graph("stacked_bar",config,noecmsmr,figmap_op_stackbar,21,"United States",
                     scenario_rename = TRUE,
                   level_scen = c("Reference","Net Zero","Net Zero+")) +
  labs(title = "") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .51),
        strip.text.y = element_text(size = 10, color = "black",face = "bold"),
        strip.text.x = element_text(size = 8, color = "black",face = "bold")) +
  bottom1
fig
ggsave(paste("output/round2/op/finals/supmat/fig",sm_fig_no,".png",sep=""), width = 11, height = 5)
```

### Non-Combustion w/ and w/o CCS

```{r}
sm_fig_no = sm_fig_no+1
sm_fig_no
df_filtered = data_final %>% filter(!model %in% c("ADAGE","FARM","gTech"))
fig <- print_graph("stacked_bar",config,df_filtered,figmap_op_stackbar,22,"United States",
                     scenario_rename = TRUE,
                   level_scen = c("Reference","Net Zero","Net Zero+")) +
  labs(title = "") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .51),
        strip.text.y = element_text(size = 10, color = "black",face = "bold"),
        strip.text.x = element_text(size = 8, color = "black",face = "bold")) +
  bottom1
fig
ggsave(paste("output/round2/op/finals/supmat/fig",sm_fig_no,".png",sep=""), width = 11, height = 5)
```

## Hydrogen Production

```{r}
sm_fig_no = sm_fig_no+1
sm_fig_no
fig <- print_graph("stacked_bar",config,data_final,figmap_op_stackbar,103,"United States",
                     scenario_rename = TRUE,
                   level_scen = c("Reference","Net Zero","Net Zero+")) +
  labs(title = "") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .51),
        strip.text.y = element_text(size = 10, color = "black",face = "bold"),
        strip.text.x = element_text(size = 10, color = "black",face = "bold")) +
  bottom1
fig
ggsave(paste("output/round2/op/finals/supmat/fig",sm_fig_no,".png",sep=""), width = 12, height = 5)
```


